---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(randomNames, tidygraph, tidyverse, igraph, ggraph)
setwd(getwd())
```


## Using Triad Census

This function counts for each triple of vertices, the number of times each of the 16 possible states occur. 

* 003  -  A,B,C, the empty graph.

* 012  -  A->B, C, the graph with a single directed edge.

* 102  -  A<->B, C, the graph with a mutual connection between two vertices.

* 021D  -  A<-B->C, the out-star.

* 021U  -  A->B<-C, the in-star.

* 021C  -  A->B->C, directed line.

* 111D  -  A<->B<-C.

* 111U  -  A<->B->C.

* 030T  -  A->B<-C, A->C.

* 030C  -  A<-B<-C, A->C.

* 201  -  A<->B<->C.

* 120D  -  A<-B->C, A<->C.

* 120U  -  A->B<-C, A<->C.

* 120C  -  A->B->C, A<->C.

* 210  -  A->B<->C, A<->C.

* 300  -  A<->B<->C, A<->C, the complete graph.

Here is an array of the names of the types 
```{r}
triad_names<- c("003 ","012 ","102 ","021D","021U","021C","111D","111U","030T","030C","201 ","120D","120U","120C","210 ","300 ")
```


### Smaller graphs
In order to explain the types, here is a selection of smaller graphs and the results of running `triad_census` on them.
```{r}
one <- graph_from_literal(A-+B, C--B)

data.frame(tri = triad_census(one), name = triad_names)
```

```{r}
two <- graph_from_literal(C-+B, B-+A, A-+C)

data.frame(tri = triad_census(two), name = triad_names)
```

```{r}
three <- graph_from_literal(A-+B, B-+A, A-+C)

data.frame(tri = triad_census(three), name = triad_names)
```

Generate list of random names for enron dataset. 
```{r}
some_names = randomNames(144, which.names="first")
```

```{r}
enron <- read_delim("dataraw/email-enron-only.mtx",delim = " ",col_names = F) %>% 
  rename("from" = X1,
         "to" = X2) %>% 
  mutate(triads_g = "default") %>% 
  mutate(name = "default") %>%
  graph_from_data_frame()

V(enron)$name <- some_names
```

```{r eval=FALSE, include=FALSE}
rm(list=ls())
```

